Reformulating phylogenetic mixed models to improve flexibility and speed

### 2019-09-10 phyloglmm_ms.tex
wrapping up for now
fixed a minor typo

We simulated 100 phylogenetic trees for each sample size (n = 25, 50, 100 and an additional n = 500 for the multi-group model) to and then modeld responses on each tree using (\ref{eq:single_glmm}, \ref{eq:multiple_glmm}). Each realization was fitted using all model variants. 
All simulation parameters are shown in Figure \ref{ssplot} and Figure \ref{msplot}. 
Correlation $\rho$ is used in place of covariances in the simulations.
Table \ref{table:platform} shows the parameters that are estimable for each platform. 
We only evaluated the goodness of fit for model fits that passed the convergence tests implemented by the package.
For Bayesian GLMM, we included only realizations where we require values of the Gelman-Rubin statistic less than the recommended threshold 1.1. Based on recent concerns about Gelman-Rubin thresholds \citep{vats2018revisiting}), we included the additional convergence criterion of effective sample size (ESS) $>1000$ for the fixed effect parameters ($\beta_{0}$ and $\beta_{1}$) for each replication \citep{vehtari2019rank}. 
For each replication, we sample two chains starting with 10000 iterations. 
%% TODO MAKE SURE THIS IS CORRECT!
We first evaluate our estimates by looking at the distribution of the estimated values (maximum likelihood estimates for non-Bayesian platforms and posterior medians for Bayesian platforms) to quantify bias and variance (i.e., quality of the point estimate).
We then compute frequentist coverage to assess the quality of the confidence intervals.
Coverage refers to the proportion of simulations in which the computed confidence intervals include the true values of parameters.
We used 95\% Wald confidence intervals for deterministic methods and quantile-based intervals (for Bayesian GLMM) to evaluate coverage.
% \bmb{clarify: did you use profile if available, otherwise Wald? You don't give any more detail in your coverage plots. [How different were Wald and profile for these results, or did you not compare them?]}
We also compare computational speed between different platforms to evaluate the efficiency of different platforms and methods.

@@

\begin{center}
\begin{figure}[H]
  \includegraphics[scale=0.7]{./figure/sstime.pdf}
  \caption{Comparison of speed for all fitting platforms.}
\label{ssplot_speed}
\end{figure}
\end{center}

\begin{center}
\begin{figure}[H]
  \includegraphics[scale=0.7]{./figure/sscoverage.pdf}
  \caption{Comparison of coverage probablility for fixed effect parameters. Models matching the simulation model (\pkg{lme4}, \pkg{brms} and \pkg{glmmTMB}) have coverage near the nominal value of 0.95. The black line shows the nominal coverage, and the grey ribbon the 95\% binomial confidence interval based on 100 simulated fits. 
%  \bmb{really picky, but maybe change grid lines or ribbon colour slightly so they're not confusable (maybe use blue ribbon with alpha=0.1?)}
  }
\label{ssplot_coverage}
\end{figure}
\end{center}

@@

\begin{center}
\begin{figure}[H]
  \includegraphics[scale=0.7]{./figure/mscoverage.pdf}
  \caption{Comparison of multi-group model coverage.}
  \label{msplot_coverage}
\end{figure}
\end{center}

@@

\newpage

@@

% \newpage
% 
% \section*{Supplements -- translation}
% 
% $1 \mid Sp_{phylo}$, $0 + X_{E} \mid Sp_{phylo}$ and $X_{E} \mid Sp_{phylo}$ means phylogenetically related have similar response,   
% 
% \begin{tabularx}{\textwidth}{|l|X|X|}
% \hline
% Formula & Statistics & Biology \\
% \hline
% $1 \mid Sp$ &
% random species intercept; variation within species in mean response across all factors &
% variation of how species respond \\
% \hline
% 
% $0 + X_{E} \mid Sp$ &
% random slope of environment factor within species; variation in coefficient within species for the environmental factor &
% variation of how species respond to the same environmental factor \\
% \hline
% 
% $1 + X_{E} \mid Sp$ &
% random slope of environmental factor within species with correlated intercept; variation in coefficient within species for the environmental factor with correlated mean response across all other factors &
% variation of how species respond in the same environmental factors and the correlation of the variation of how they respond in general \\
% \hline
% 
% $1 \mid Site:Sp $ &
% random variation in intercept among species within sites &
% variation of how species respond within sites \\
% \hline
% 
% $1 | Sp_{Phylo} $ &
% variation among species in mean response across all factors demonstrate phylogenetic signal &
% phylogenetically related species respond similarly \\
% \hline
% 
% $0 + X_{E} \mid Sp_{Phylo}$ &
% variation among species for environmental factors demonstrate phylogenetic signal &
% phylogenetically related species respond similar (share common response) to the same environmental factor \\
% \hline
% \end{tabularx}
%             
%             
%             
%             
%             
%

>>>>

We simulated 100 phylogenetic trees for each sample size (n = 25, 50, 100 and an additional n = 500 for the multi-group model) to and then modelled responses on each tree using (\ref{eq:single_glmm}, \ref{eq:multiple_glmm}). Each realization was fitted using all model variants. 
All simulation parameters are shown in Figure \ref{ssplot} and Figure \ref{msplot}. 
Correlation $\rho$ is used in place of covariances in the simulations.
Table \ref{table:platform} shows the parameters that are estimable for each platform. 
We only evaluated the goodness of fit for model fits that passed the convergence tests implemented by the package.
For Bayesian GLMM, we included only realizations where we require values of the Gelman-Rubin statistic less than the recommended threshold 1.1. Based on recent concerns about Gelman-Rubin thresholds \citep{vats2018revisiting}), we included the additional convergence criterion of effective sample size (ESS) $>1000$ for the fixed effect parameters ($\beta_{0}$ and $\beta_{1}$) for each replication \citep{vehtari2019rank}. 
For each replication, we sample two chains starting with 10000 iterations. 
%% TODO MAKE SURE THIS IS CORRECT!
We first evaluate our estimates by looking at the distribution of the estimated values (maximum likelihood estimates for non-Bayesian platforms and posterior medians for Bayesian platforms) to quantify bias and variance (i.e., quality of the point estimate).
We then compute frequentist coverage to assess the quality of the confidence intervals.
Coverage refers to the proportion of simulations in which the computed confidence intervals include the true values of parameters.
We used 95\% Wald confidence intervals for deterministic methods and quantile-based intervals (for Bayesian GLMM) to evaluate coverage.
% \bmb{clarify: did you use profile if available, otherwise Wald? You don't give any more detail in your coverage plots. [How different were Wald and profile for these results, or did you not compare them?]}
We also compare computational speed between different platforms to evaluate the efficiency of different platforms and methods.

@@

\begin{center}
\begin{figure}[H]
  \includegraphics[scale=0.7]{./figure/sstime.pdf}
  \caption{Comparison of single-group model computational speed.}
\label{ssplot_speed}
\end{figure}
\end{center}

\begin{center}
\begin{figure}[H]
  \includegraphics[scale=0.7]{./figure/sscoverage.pdf}
  \caption{Comparison of coverage probablility for fixed effect parameters. Models matching the simulation model (\pkg{lme4}, \pkg{glmmTMB} and pkg{brms}) have coverage near the nominal value of 0.95. The black line shows the nominal coverage, and the blue ribbon the 95\% binomial confidence interval based on 100 simulated fits. 
%  \bmb{really picky, but maybe change grid lines or ribbon colour slightly so they're not confusable (maybe use blue ribbon with alpha=0.1?)}
  }
\label{ssplot_coverage}
\end{figure}
\end{center}

@@

\begin{center}
\begin{figure}[H]
  \includegraphics[scale=0.5]{./figure/mscoverage.pdf}
  \caption{Comparison of multi-group model coverage.}
  \label{msplot_coverage}
\end{figure}
\end{center}

@@

@@

% \newpage
% 
% \section*{Supplements -- translation}
% 
% $1 \mid Sp_{phylo}$, $0 + X_{E} \mid Sp_{phylo}$ and $X_{E} \mid Sp_{phylo}$ means phylogenetically related have similar response,   
% 
% \begin{tabularx}{\textwidth}{|l|X|X|}
% \hline
% Formula & Statistics & Biology \\
% \hline
% $1 \mid Sp$ &
% random species intercept; variation within species in mean response across all factors &
% variation of how species respond \\
% \hline
% 
% $0 + X_{E} \mid Sp$ &
% random slope of environment factor within species; variation in coefficient within species for the environmental factor &
% variation of how species respond to the same environmental factor \\
% \hline
% 
% $1 + X_{E} \mid Sp$ &
% random slope of environmental factor within species with correlated intercept; variation in coefficient within species for the environmental factor with correlated mean response across all other factors &
% variation of how species respond in the same environmental factors and the correlation of the variation of how they respond in general \\
% \hline
% 
% $1 \mid Site:Sp $ &
% random variation in intercept among species within sites &
% variation of how species respond within sites \\
% \hline
% 
% $1 | Sp_{Phylo} $ &
% variation among species in mean response across all factors demonstrate phylogenetic signal &
% phylogenetically related species respond similarly \\
% \hline
% 
% $0 + X_{E} \mid Sp_{Phylo}$ &
% variation among species for environmental factors demonstrate phylogenetic signal &
% phylogenetically related species respond similar (share common response) to the same environmental factor \\
% \hline
% \end{tabularx}
%             
%             
%             
%             
%             
%                                                               
\singlespacing
